---
title: "Tidying"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, cache = FALSE}
source(here::here("code", "setup.R"))
```

# Introduction {.unnumbered}

After the scraping process we have a the data about songs played during the
Requestival in a nice tabular format. Let's load it up and see what it looks
like.

```{r load}
requestival <- read_tsv(
    PATHS$scraped,
    col_types = cols(
        .default = col_character()
    )
)
```

This isn't too messy at the moment but there are some things we could do to tidy
it up. Let's work through each of the columns and see if we need to do anything
to them.

# File

The first column contains the name of the file the song was scraped from. The
file names have the form `requestival_X`, where `X` is a day in May 2020. We
don't really care about the file name but we **do** care about the day the songs
were played so let's take that part and create a new column.

```{r add-day}
requestival <- requestival %>%
    mutate(Day = str_remove(File, "requestival_")) %>%
    mutate(Day = paste0("2020-05-", Day))

requestival
```

# Time

The time each song was played is currently a string with the form `hh:mmpp`
(where `pp` is `am` or `pm`). It would be better to have this has a time object
so let's do that conversion.

```{r parse-time}
requestival <- requestival %>%
    mutate(Time = parse_time(Time, "%I:%M%p"))

requestival
```

Now all our times are `hms` objects and that have the form `HH:MM:ss`. Our days
and times are now in nice formats let's combine them into a single datetime.
This is pretty easy to do, we just covert our day string to a `datetime` object
and then add on our time.

```{r datetime}
requestival <- requestival %>%
    mutate(DateTime = as_datetime(Day, tz = "Europe/Berlin") + Time)

requestival
```

You might notice that the times are in the CEST timezone. This is because the
HTML pages were downloaded in Europe and the Triple J website is clever enough
to display times using your local timezone. It doesn't really make sense to use
European times for a dataset from an Australian radio station so let's convert
them to the Australian east coast timezone that Triple J broadcasts in.

```{r convert-tz}
requestival <- requestival %>%
    mutate(DateTime = with_tz(DateTime, "Australia/Sydney"))

requestival
```

# Song

The song name column is made up of strings. This is probably what we want but
let's do a quick check to see if we have any weird characters left over from the
HTML.

```{r song-chars}
song_chars <- requestival$Song %>%
    str_extract_all(boundary("character")) %>% 
    unlist() %>%
    sort() %>%
    unique()
```

These are the unique characters we have in the song names:
`r glue_collapse(song_chars, sep = ", ")`

Most of these letters in the standard Latin alphabet, numerals or standard
punctuation but there are also a few different kinds of white space, some
accented letters and what I think are Korean characters.

Let's make a table of these characters with their Unicode descriptions to try
and work out what they are.

```{r song-chars-table}
song_chars %>%
    map_int(utf8ToInt) %>%
    Unicode::u_char_info() %>%
    mutate(Character = song_chars) %>%
    select(Character, Code, Name)
```

I have quickly checked some of these and it looks like they are correctly part
of song names so I am going to leave them. I will replace the non-breaking
spaces with regular spaces though.

```{r song-spaces}
requestival <- requestival %>%
    mutate(Song = str_replace(Song, "\u00a0", " "))
```

# Artist

We can do a similar thing to have a look at the characters in the artist field.

```{r artist-chars}
artist_chars <- requestival$Artist %>%
    str_extract_all(boundary("character")) %>% 
    unlist() %>%
    sort() %>%
    unique()

artist_chars %>%
    map_int(function(.x) {utf8ToInt(.x)[1]}) %>%
    Unicode::u_char_info() %>%
    mutate(Character = artist_chars) %>%
    select(Character, Code, Name)
```

This looks ok so I don't think we need to do anything here.

# Release

Let's have a look at the characters in the release field as well.

```{r release-chars}
release_chars <- requestival$Release %>%
    str_extract_all(boundary("character")) %>% 
    unlist() %>%
    sort() %>%
    unique()

release_chars %>%
    map_int(function(.x) {utf8ToInt(.x)[1]}) %>%
    Unicode::u_char_info() %>%
    mutate(Character = release_chars) %>%
    select(Character, Code, Name)
```

There are some more unusual characters here but they also look to be correct so
we will leave them for now.

# Links

The remaining fields that we scraped are links to places where you can listen
to each song. We probably don't want to mess with these too much but let's take
a look.

## YouTube

The YouTube URLs look something like this: `r requestival$YouTube[1]`

This is just a link a YouTube search results page for a query made up of the
artist and song name. It might be useful to be able to access the query part so
let's make that into a new column.

```{r youtube-query}
requestival <- requestival %>%
    mutate(YouTubeQuery = str_remove(YouTube, "^.*="))

requestival
```

## Spotify

The Spotify URLs are similar: `r requestival$Spotify[1]`

We can make a similar column for the Spotify query.

```{r spotify-query}
requestival <- requestival %>%
    mutate(SpotifyQuery = str_remove(Spotify, "^.*/results/"))

requestival
```

## Unearthed

The Unearthed links are a bit different: `r requestival$Unearthed[3]`

These are direct links to artist pages on the Unearthed website rather than
searches. There probably isn't much useful information in these but it would be
nice to know which artists are on Unearthed so let's make a boolean column to
represent this.

```{r is-unearthed}
requestival <- requestival %>%
    mutate(IsUnearthed = !is.na(Unearthed))

requestival
```

# Selecting columns

Now that we have tidied up the individual columns we can tidy up the table as a
whole. For example the "File" field isn't much use now that we have extracted the
date information and the "Day", "Time" and "DateTime" columns are a bit
redundant. The times are also inconsistent because we have converted DateTime to
the Australian time zone but the others are still in European time. We might
also like to rearrange and rename some of the columns.

```{r select}
requestival <- requestival %>%
    select(
        DateTime,
        Song,
        Artist,
        Release,
        IsUnearthed,
        UnearthedURL = Unearthed,
        SpotifyQuery,
        SpotifyURL   = Spotify,
        YouTubeQuery,
        YouTubeURL   = YouTube
    ) %>%
    arrange(DateTime)

requestival
```

# Filtering

The final step in our tidying is to filter out some of the songs.
The Requestival officially runs between 6 am and 9 pm from Monday 25 May to
Sunday 31 May. On the final day it actually ends a bit earlier at 6 pm. We are
only interested in songs played during this period so we need to filter out
other times. There is probably a cleaner way to do this but I'm just going to
use a big ol' logical statement.

```{r filter}
tz <- "Australia/Sydney" # This is just to save space
times <- tribble(
                                 ~Start,                                ~End,
    ymd_hm("2020-05-25 06:00", tz = tz), ymd_hm("2020-05-25 21:00", tz = tz),
    ymd_hm("2020-05-26 06:00", tz = tz), ymd_hm("2020-05-26 21:00", tz = tz),
    ymd_hm("2020-05-27 06:00", tz = tz), ymd_hm("2020-05-27 21:00", tz = tz),
    ymd_hm("2020-05-28 06:00", tz = tz), ymd_hm("2020-05-28 21:00", tz = tz),
    ymd_hm("2020-05-29 06:00", tz = tz), ymd_hm("2020-05-29 21:00", tz = tz),
    ymd_hm("2020-05-30 06:00", tz = tz), ymd_hm("2020-05-30 21:00", tz = tz),
    ymd_hm("2020-05-31 06:00", tz = tz), ymd_hm("2020-05-31 18:00", tz = tz)
)

requestival <- requestival %>%
    filter(
        (DateTime >= times$Start[1] & DateTime <= times$End[1]) |
            (DateTime >= times$Start[2] & DateTime <= times$End[2]) |
            (DateTime >= times$Start[3] & DateTime <= times$End[3]) |
            (DateTime >= times$Start[4] & DateTime <= times$End[4]) |
            (DateTime >= times$Start[5] & DateTime <= times$End[5]) |
            (DateTime >= times$Start[6] & DateTime <= times$End[6]) |
            (DateTime >= times$Start[7] & DateTime <= times$End[7])
    )
```

It would be easy to miss this up let's make a quick plot of the play times to
check what songs we have kept.

```{r plot-times}
ggplot(requestival, aes(x = DateTime, y = 0)) +
    geom_jitter(width = 0) +
    geom_vline(data = times, aes(xintercept = Start), colour = "red") +
    geom_vline(data = times, aes(xintercept = End), colour = "blue") +
    theme_minimal()
```

That looks pretty good 🎉!

The final dataset has **`r ncol(requestival)`** columns and
**`r nrow(requestival)`** rows. Let's save it to use for future analysis.

```{r save}
write_tsv(requestival, PATHS$tidied)

requestival
```
