---
title: "Augmentation"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, cache = FALSE}
source(here::here("code", "setup.R"))
```

# Introduction {.unnumbered}

The tidy we scraped from the HTML pages is now in a nice clean and tidy format
but maybe there is some other information we can add to it from other sources?

```{r load}
requestival <- read_tsv(
    PATHS$tidied,
    col_types = cols(
        DateTime     = col_datetime(format = ""),
        Song         = col_character(),
        Artist       = col_character(),
        Release      = col_character(),
        IsUnearthed  = col_logical(),
        UnearthedURL = col_character(),
        SpotifyQuery = col_character(),
        SpotifyURL   = col_character(),
        YouTubeQuery = col_character(),
        YouTubeURL   = col_character()
    )
) %>%
    mutate(DateTime = with_tz(DateTime, "Australia/Sydney"))
```

# Spotify

We already have search queries for Spotify so maybe we can pull some useful
information from their database? Luckily they have a nice API that we can access
with the **{spotifyr}** package. See the
[**{spotifyr}** documentation][spotifyr] for more details about how this works.

```{r spotifyr}
access_token <- get_spotify_access_token(
    client_id     = SPOTIFY_CLIENT_ID,
    client_secret = SPOTIFY_CLIENT_SECRET
)
```

## Track IDs

To get more information about each song we need track IDs which we don't
currently have. The queries we extracted from the Spotify links give use a clue
about how to format the search terms but unfortunately we can't use them 
directly 😿. Because they were designed for the web spaces and other characters
have been replaced with codes (e.g. `%20` for space) but the **{spotifyr}**
search function prefers the standard characters. It's not hard to construct
the queries from the information we have though.

For some songs we will get multiple results. Here we try and handle that by
first search for the song, artist and release. If there are any results we
return the first one. If not we try a simpler search with just the song and 
artist, and again return the first result.

```{r track-ids}
get_id <- function(Song, Artist, Release, ...) {
    
    # Search for artist, track, album
    query <- glue('artist:"{Artist}" track:"{Song}" album: "{Release}"')
    
    results <- search_spotify(
        query,
        type          = "track",
        market        = "AU",
        authorization = access_token
    ) %>%
        filter(album.release_date_precision == "day")
    
    # Return the earliest result if there are any
    if (nrow(results) > 0) {
        return(results$id[1])
    }
    
    # Otherwise search for just artist and track
    query <- glue('artist:"{Artist}" track:"{Song}"')
    
    results <- search_spotify(
        query,
        type          = "track",
        market        = "AU",
        authorization = access_token
    ) %>%
        filter(album.release_date_precision == "day")
    
    # Return the first result or NA
    if (nrow(results) > 0) {
        return(results$id[1])
    } else {
        return(NA)
    }
}

requestival <- requestival %>%
    mutate(SpotifyID = pmap_chr(requestival, get_id)) %>%
    mutate(HasSpotify = !is.na(SpotifyID))
```

The track IDs are character strings that look something like this:
`r requestival$SpotifyID[1]`. For our **`r nrow(requestival)`** songs we were
able to find **`r sum(!is.na(requestival$SpotifyID))`** IDs. These IDs aren't
very interesting by themselves but we can use them to retrieve other information 
from the Spotify database.

## Track information

Now that we have track IDs we can get some more information about each of the
Requestival songs.

```{r track-info}
tracks <- requestival %>%
    filter(HasSpotify) %>%
    pull(SpotifyID) %>%
    map_dfr(function(.id) {
        track_info <- get_track(
            .id,
            market        = "AU",
            authorization = access_token
        )
        tibble(
            SpotifyID  = track_info$id,
            AlbumDate  = track_info$album$release_date,
            Duration   = track_info$duration_ms,
            Explicit   = track_info$explicit,
            Popularity = track_info$popularity
        )
    }) %>%
    mutate(
        Duration  = Duration / 1000,
        AlbumDate = ymd(AlbumDate)
    )

tracks
```

A lot of the track information isn't useful or is redundant with what we already
have but I have picked out a few things that might be interesting: the track
duration (in seconds), whether it is explicit or not and the Spotify popularity
score.

## Audio features

We can also use the Spotify API to retrieve some information about the audio
features of each song.

```{r audio-features}
audio <- requestival %>%
    filter(HasSpotify) %>%
    pull(SpotifyID) %>%
    map_dfr(function(.id) {
        audio_features <- get_track_audio_features(
            .id,
            authorization = access_token
        )
    }) %>%
    mutate(IsMajor = mode == 1) %>%
    select(
        SpotifyID    = id,
        IsMajor,
        Loudness     = loudness,
        Tempo        = tempo,
        Valence      = valence,
        Energy       = energy,
        Danceability = danceability,
        Speechiness  = speechiness,
        Acousticness = acousticness,
        Liveness     = liveness
    )

audio
```

Detail about the audio features can be found [here][audio-features] but in
general they try to provide a numeric description of the track. Some things like
loudness, tempo or whether the track is in a major key are simply calculated
from the track but others such as "danceability" or "acousticness" are more
abstract summaries of what the track sounds like.

Let's join the information we have downloaded from Spotify to our data scraped
from the HTML pages and save it for analysis.

```{r save}
requestival <- requestival %>%
    left_join(tracks, by = "SpotifyID") %>%
    left_join(audio, by = "SpotifyID") %>%
    distinct()

write_tsv(requestival, PATHS$augmented)

requestival
```

[spotifyr]: https://www.rcharlie.com/spotifyr/ "spotifyr website"
[audio-features]: https://developer.spotify.com/documentation/web-api/reference/tracks/get-several-audio-features/ "Get several audio features"
