---
title: "Exploration"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, cache = FALSE}
source(here::here("code", "setup.R"))
```

# Introduction {.unnumbered}

In this document we are going to do some basic exploration of the complete
augmented dataset. We will work through each column make some basics plots and
and summaries. This should have to give us a better sense of the data but might
also expose any mistakes we made during the pre-processing stages. 

```{r load}
requestival <- read_tsv(
    PATHS$augmented,
    col_types = cols(
        .default     = col_double(),
        DateTime     = col_datetime(format = ""),
        Song         = col_character(),
        Artist       = col_character(),
        Release      = col_character(),
        IsUnearthed  = col_logical(),
        UnearthedURL = col_character(),
        SpotifyQuery = col_character(),
        SpotifyURL   = col_character(),
        YouTubeQuery = col_character(),
        YouTubeURL   = col_character(),
        SpotifyID    = col_character(),
        HasSpotify   = col_logical(),
        AlbumDate    = col_date(format = ""),
        Explicit     = col_logical(),
        IsMajor      = col_logical()
   )
) %>%
    mutate(DateTime = with_tz(DateTime, "Australia/Sydney"))
```

The dataset has **`r nrow(requestival)`** rows and **`r ncol(requestival)`** 
columns.

# Triple J features

Let's start with the features we scraped from the HTML files.

## DateTime

When were the songs played?

```{r DateTime}
ggplot(requestival, aes(x = DateTime)) +
    geom_histogram(bins = 200)
```

## Song

There are **`r length(unique(requestival$Song))`** unique songs. How many times
was each song played?

```{r Song-count}
song_counts <- requestival %>%
    group_by(Song, Artist) %>%
    count(name = "PlayCount")

ggplot(song_counts, aes(x = PlayCount)) +
    geom_histogram()
```

Which songs were played more than once?

```{r Song-multiple}
song_counts %>%
    filter(PlayCount > 1) %>%
    arrange(-PlayCount)
```

## Artist

There are **`r length(unique(requestival$Artist))`** unique artists. How many 
times was each artist played?

```{r Artist-count}
artist_counts <- requestival %>%
    group_by(Artist) %>%
    count(name = "PlayCount")

ggplot(artist_counts, aes(x = PlayCount)) +
    geom_histogram()
```

Which artists were played more than once?

```{r Artist-multiple}
artist_counts %>%
    filter(PlayCount > 1) %>%
    arrange(-PlayCount)
```

## Release

There are **`r length(unique(requestival$Release))`** unique releases. How many 
times was each release played?

```{r Release-count}
release_counts <- requestival %>%
    group_by(Release) %>%
    count(name = "PlayCount")

ggplot(release_counts, aes(x = PlayCount)) +
    geom_histogram()
```

Which releases were played more than once?

```{r Release-multiple}
release_counts %>%
    filter(PlayCount > 1) %>%
    arrange(-PlayCount)
```

Which songs do not have an associated release?

```{r Release-NA}
requestival %>%
    filter(is.na(Release)) %>%
    select(DateTime, Song, Artist)
```

This seems weird but I have checked them and this information is missing from
the original HTML pages. It's only a few songs so I'm not going to try and fix
it.

## Unearthed

How many songs are on Unearthed?

```{r Unearthed}
ggplot(requestival, aes(x = IsUnearthed)) +
    geom_bar()
```

# Spotify

Now let's looks at the fields we downloaded from Spotify. How many songs did we
find Spotify track IDs for?

```{r HasSpotify}
ggplot(requestival, aes(x = HasSpotify)) +
    geom_bar()
```

For the rest of this section we will only look at the songs with Spotify
information.

```{r filter-spotify}
requestival_spotify <- filter(requestival, HasSpotify)
```

## Album date

When we the songs released? This is the album release date so may not be the
earliest song release depending on which album we got from Spotify.

```{r AlbumDate}
ggplot(requestival_spotify, aes(x = AlbumDate)) +
    geom_histogram()
```

The five most recent songs are:

```{r AlbumDate-recent}
requestival_spotify %>%
    arrange(desc(AlbumDate)) %>%
    top_n(5, AlbumDate) %>%
    select(Song, Artist, Release, AlbumDate)
```

The five oldest songs are:

```{r AlbumDate-old}
requestival_spotify %>%
    arrange(AlbumDate) %>%
    top_n(-5, AlbumDate) %>%
    select(Song, Artist, Release, AlbumDate)
```

## Duration

How long are the songs?

```{r Duration}
ggplot(requestival_spotify, aes(x = Duration)) +
    geom_histogram(bins = 100) +
    scale_x_time()
```

The five longest songs are:

```{r Duration-long}
requestival_spotify %>%
    arrange(-Duration) %>%
    top_n(5, Duration) %>%
    select(Song, Artist, Duration)
```

The five shortest songs are:

```{r Duration-short}
requestival_spotify %>%
    arrange(Duration) %>%
    top_n(-5, Duration) %>%
    select(Song, Artist, Duration)
```

## Explicit

How many songs are explicit?

```{r Explicit}
ggplot(requestival_spotify, aes(x = Explicit)) +
    geom_bar()
```

## Popularity

What is the distribution of popularity scores?

```{r Popularity}
ggplot(requestival_spotify, aes(x = Popularity)) +
    geom_bar()
```

The five most "popular" songs are:

```{r Popularity-high}
requestival_spotify %>%
    arrange(-Popularity) %>%
    top_n(5, Popularity) %>%
    select(Song, Artist, Popularity)
```

The five least "popular" songs are:

```{r Popularity-low}
requestival_spotify %>%
    arrange(Popularity) %>%
    top_n(-5, Popularity) %>%
    select(Song, Artist, Popularity)
```

## IsMajor

How many songs are in a major key?

```{r IsMajor}
ggplot(requestival_spotify, aes(x = IsMajor)) +
    geom_bar()
```

## Loudness

What is the distribution of loudness?

```{r Loudness}
ggplot(requestival_spotify, aes(x = Loudness)) +
    geom_histogram(bins = 100)
```

The five loudest songs are:

```{r Loudness-loud}
requestival_spotify %>%
    arrange(-Loudness) %>%
    top_n(5, Loudness) %>%
    select(Song, Artist, Loudness)
```

The five quietest songs are:

```{r Loudness-quiet}
requestival_spotify %>%
    arrange(Loudness) %>%
    top_n(-5, Loudness) %>%
    select(Song, Artist, Loudness)
```

## Tempo

What speed are the songs?

```{r Temp}
ggplot(requestival_spotify, aes(x = Tempo)) +
    geom_histogram(bins = 100)
```

The five fastest songs are:

```{r Temp-fast}
requestival_spotify %>%
    arrange(-Tempo) %>%
    top_n(5, Tempo) %>%
    select(Song, Artist, Tempo)
```

The five slowest songs are:

```{r Tempo-slow}
requestival_spotify %>%
    arrange(Tempo) %>%
    top_n(-5, Tempo) %>%
    select(Song, Artist, Tempo)
```

## Valence

What is the distribution of valence? This is score from zero to one where one
is positive and zero is negative.

```{r Valence}
ggplot(requestival_spotify, aes(x = Valence)) +
    geom_histogram(bins = 100)
```

The five most positive songs are:

```{r Valence-high}
requestival_spotify %>%
    arrange(-Valence) %>%
    top_n(5, Valence) %>%
    select(Song, Artist, Valence)
```

The five most negative songs are:

```{r Valence-low}
requestival_spotify %>%
    arrange(Valence) %>%
    top_n(-5, Valence) %>%
    select(Song, Artist, Valence)
```

## Energy

What is the distribution of energy?

```{r Energy}
ggplot(requestival_spotify, aes(x = Energy)) +
    geom_histogram(bins = 100)
```

The five most energetic songs are:

```{r Energy-high}
requestival_spotify %>%
    arrange(-Energy) %>%
    top_n(5, Energy) %>%
    select(Song, Artist, Energy)
```

The five least energetic songs are:

```{r Energy-low}
requestival_spotify %>%
    arrange(Energy) %>%
    top_n(-5, Energy) %>%
    select(Song, Artist, Energy)
```

## Danceability

What is the distribution of danceability?

```{r Danceability}
ggplot(requestival_spotify, aes(x = Danceability)) +
    geom_histogram(bins = 100)
```

The five most danceable songs are:

```{r Danceability-high}
requestival_spotify %>%
    arrange(-Danceability) %>%
    top_n(5, Danceability) %>%
    select(Song, Artist, Danceability)
```

The five least danceable songs are:

```{r Danceability-low}
requestival_spotify %>%
    arrange(Danceability) %>%
    top_n(-5, Danceability) %>%
    select(Song, Artist, Danceability)
```

## Speechiness

What is the distribution of speechiness?

```{r Speechiness}
ggplot(requestival_spotify, aes(x = Speechiness)) +
    geom_histogram(bins = 100)
```

The five most speechy songs are:

```{r Speechiness-high}
requestival_spotify %>%
    arrange(-Speechiness) %>%
    top_n(5, Speechiness) %>%
    select(Song, Artist, Speechiness)
```

The five least speechy songs are:

```{r Speechiness-low}
requestival_spotify %>%
    arrange(Speechiness) %>%
    top_n(-5, Speechiness) %>%
    select(Song, Artist, Speechiness)
```

## Acousticness

What is the distribution of acousticness?

```{r Acousticness}
ggplot(requestival_spotify, aes(x = Acousticness)) +
    geom_histogram(bins = 100)
```

The five most acoustic songs are:

```{r Acousticness-high}
requestival_spotify %>%
    arrange(-Acousticness) %>%
    top_n(5, Acousticness) %>%
    select(Song, Artist, Acousticness)
```

The five least acoustic are:

```{r Acousticness-low}
requestival_spotify %>%
    arrange(Acousticness) %>%
    top_n(-5, Acousticness) %>%
    select(Song, Artist, Acousticness)
```

## Liveness

What is the distribution of liveness?

```{r Liveness}
ggplot(requestival_spotify, aes(x = Liveness)) +
    geom_histogram(bins = 100)
```

The five most live songs are:

```{r Liveness-high}
requestival_spotify %>%
    arrange(-Liveness) %>%
    top_n(5, Liveness) %>%
    select(Song, Artist, Liveness)
```

The five least live songs are:

```{r Liveness-low}
requestival_spotify %>%
    arrange(Liveness) %>%
    top_n(-5, Liveness) %>%
    select(Song, Artist, Liveness)
```
