---
title: "Embedding"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, cache = FALSE}
source(here::here("code", "setup.R"))
```

# Introduction {.unnumbered}

After checking the individual variables the first thing I like to do with a new
dataset is to perform some dimensionality reduction to get an overiew of it as a
whole.

```{r load}
requestival <- read_tsv(
    PATHS$augmented,
    col_types = cols(
        .default     = col_double(),
        DateTime     = col_datetime(format = ""),
        Song         = col_character(),
        Artist       = col_character(),
        Release      = col_character(),
        IsUnearthed  = col_logical(),
        UnearthedURL = col_character(),
        SpotifyQuery = col_character(),
        SpotifyURL   = col_character(),
        YouTubeQuery = col_character(),
        YouTubeURL   = col_character(),
        SpotifyID    = col_character(),
        HasSpotify   = col_logical(),
        AlbumDate    = col_date(format = ""),
        Explicit     = col_logical(),
        IsMajor      = col_logical()
)
) %>%
    mutate(DateTime = with_tz(DateTime, "Australia/Sydney"))
```

# PCA

Let's perform a PCA of the dataset. We need numeric values so we can only really
do this on the songs that we have Spotify information for.

```{r pca-features}
features <- c("Duration", "Loudness", "Tempo", "Valence", "Energy", 
              "Danceability", "Speechiness", "Acousticness", "Liveness")

features_mat <- requestival %>%
    filter(HasSpotify) %>%
    select(SpotifyID, !!features) %>%
    distinct() %>%
    column_to_rownames("SpotifyID") %>%
    as.matrix() %>%
    scale()
```

I have selected the features that describe the basic characteristics of each
song. These are:
`r glue_collapse(glue("**{features}**"), sep = ", ", last = " and ")`. Most of
these features are scores between zero and one but a few have very different
ranges so I have centered and scaled each of them.

```{r pca}
pca <- prcomp(features_mat)
```

Let's have a look at the loadings for each principle component.

```{r pca-loadings}
feature_loadings <- pca$rotation %>%
    as.data.frame() %>%
    rownames_to_column("Feature") %>%
    mutate(Feature = factor(Feature, levels = features)) %>%
    pivot_longer(-Feature, names_to = "PC", values_to = "Loading")

ggplot(feature_loadings,
       aes(x = fct_rev(Feature), y = Loading, fill = Feature)) +
    geom_col() +
    geom_hline(yintercept = 0) +
    coord_flip() +
    scale_fill_brewer(palette = "Set1", guide = FALSE) +
    facet_wrap(~ PC) +
    labs(
        title = "PC loadings",
        x     = NULL
    ) +
    theme(
        strip.text       = element_text(colour = "white"),
        strip.background = element_rect(fill = "black"),
        panel.border     = element_rect(fill = NA)
    )
```

Here is the scatter plot of the first two components.

```{r pca-scatter}
pca_points <- pca$x %>%
    as.data.frame() %>%
    rownames_to_column("SpotifyID")

ggplot(pca_points, aes(x = PC1, y = PC2)) +
    geom_point()
```

There's not a lot of separation between points here. Perhaps thing will be
clearer if we do a non-linear embedding based on the PCA?

# t-SNE

Let's see what a t-SNE plot looks like.

```{r tsne}
tsne <- Rtsne(pca$x, pca = FALSE)

tsne_points <- tsne$Y %>%
    as.data.frame() %>%
    set_names(c("TSNE1", "TSNE2")) %>%
    mutate(SpotifyID = pca_points$SpotifyID)

ggplot(tsne_points, aes(x = TSNE1, y = TSNE2)) +
    geom_point()
```

That maybe does a better job of showing the structure in the data? What if we
overlay some of the information we have.

## Features {.tabset}

```{r tsne-features, results = "hide"}
requestival <- left_join(requestival, tsne_points, by = "SpotifyID")

plot_data <- requestival %>%
    filter(HasSpotify) %>%
    select(-DateTime) %>%
    distinct(SpotifyID, .keep_all = TRUE)

plot_features <- c("IsUnearthed", "AlbumDate", "Duration", "Explicit",
                   "Popularity", "IsMajor", "Loudness", "Tempo", "Valence",
                   "Energy", "Danceability", "Speechiness", "Acousticness",
                   "Liveness")

src_list <- map(plot_features, function(.feature) {
    if (is.logical(plot_data[[.feature]][1])) {
        colour_scale <- 'scale_colour_manual(values = c("grey", "red"))'
    } else if (is.Date(plot_data[[.feature]][1])) {
        colour_scale <- 'scale_colour_date()'
    } else {
        colour_scale <- 'scale_colour_viridis_c()'
    }
    knit_expand(text = c(
        "### {{.feature}} {.unnumbered}",
        "```{r}",
        "ggplot(plot_data) +",
        "aes(x = TSNE1, y = TSNE2, colour = {{.feature}}) +",
        "geom_point(size = 2) +",
        colour_scale,
        "```",
        ""
    ))
})

out <- knit_child(text = unlist(src_list), options = list(cache = FALSE))
```

`r out`

requestival %>% filter(HasSpotify) %>% mutate(Day = day(DateTime), Time = (60 * hour(DateTime) + minute(DateTime)) * 60) %>% select(Day, Time, Danceability, Speechiness, Acousticness, Liveness) %>% pivot_longer(c(-Time, -Day), names_to = "Feature", values_to = "Score") -> pp

ggplot(pp, aes(x = Time, y = Score, colour = Feature, fill = Feature)) + geom_point(alpha = 0.25, size = 1) + geom_smooth() + facet_grid(Day ~ Feature, scales = "free") + scale_x_time() + scale_colour_brewer(palette = "Set1", guide = FALSE) + scale_fill_brewer(palette = "Set1", guide = FALSE)
